---
title: "First Markdown"
author: "Sophie Kogut"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

$a + b$
$$a + b$$
$$\alpha + \beta$$
$$S_a$$
$$S_{a+b_{\alpha + \beta}}$$
$$Z^{a + b}$$
$$P(occurence of species A) = Z_A$$
$$P(\mbox{occurence of species A}) = Z_A$$

## R Markdown

first line.   
second line.   
third line.  
space space space

```
plain text and nothing else.
$$\alpha + \beta$$
```

Use a single back tick ` for plain text in a sentence `.

The value of $\pi$ is `r pi`.   

```{r echo=TRUE, eval=FALSE}
# comment!
runif(10)
```




```{bash}
Using shell commands
pwd
```
```{r}

```


When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
