---
title: "Homework5"
author: "Sophie Kogut"
date: "2/12/2020"
output: html_document
---
## Working with matrices, lists, and data frames
1. Assign to the variable n_dims a single random integer between 3 and 10.
- Create a vector of consecutive integers from 1 to n_dims^2.
- Use the sample function to randomly reshuffle these values.
- create a square matrix with these elements.
- print out the matrix.
- find a function in r to transpose the matrix.
- print it out again and note how it has changed.
- calculate the sum and the mean of the elements in the first row and the last row.
- read about the eigen() function and use it on your matrix
- look carefully at the elements of values and vectors. What kind of numbers are these?
- dig in with the typeof() function to figure out their type.
- if have set your code up properly, you should be able to re-run it and create a matrix of different size because n_dims will change.

```{r}
n_dims <- sample(3:10, 1)
print(n_dims)
my_vec <- seq(from= 1, to= (n_dims^2))
print(my_vec)
reshuffle<- sample(my_vec, length(my_vec), replace = FALSE)
# can also use sample(my_vec)
print(reshuffle)
my_mat<- matrix(nrow= n_dims, ncol= n_dims, data= reshuffle)
print(my_mat)
my_mat <- t(my_mat)
print(my_mat)
# The transpose function switches the rows and columns of the matrix.
sum(my_mat[1,]) # sum of first row
mean(my_mat[1,]) # mean of first row
sum(my_mat[n_dims,]) # sum of last row
mean(my_mat[n_dims,]) # mean of last row
? eigen
eigen_m<- eigen(my_mat)
typeof(eigen_m$values)
typeof(eigen_m$vectors)
# The values and vectors are of type "complex", indicating they defined by the imaginery number i. 
```


2. Create a list with the following named elements:
- my_matrix, which is a 4 x 4 matrix filled with random uniform values
- my_logical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.
- my_letters, which is a 26-element vector of all the lower-case letters in random order.
- create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
- use the typeof() function to confirm the underlying data types of each component in this list
- combine the underlying elements from the new list into a single atomic vector with the c() function.
- what is the data type of this vector?

```{r}
my_list <- list(my_matrix= matrix(nrow=4, ncol=4, data= runif(16)),
             my_logical<- runif(100) > 0.5, my_letters= sample(letters[1:26]))
print(my_list)

list2 <- list(my_list$my_matrix[2,2], my_list[[2]][2], my_list$my_letters[2])
print(list2)

typeof(list2[[1]])
typeof(list2[[2]])
typeof(list2[[3]])

vec<- c(list2[[1]],list2[[2]], list2[[3]])
typeof(vec)

```


3. Create a data frame with two variables (= columns) and 26 cases (= rows).
- call the first variable my_unis and fill it with a random uniform value from 0 to 10
- call the second variable my_letters and fill it with 26 capital letters in random order.
- for the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.
- for the first variable, write a single line of R code to identify which rows have the missing values.
- for the second variable, sort it in alphabetical order
- calculate the column mean for the first variable.
```{r}
dframe<- data.frame("my_unis"= runif(26, min=0, max=10), "my_letters"= sample(LETTERS[1:26]) )
print(dframe)

dframe[sample(nrow(dframe),4, replace=FALSE),1] <- NA

which(is.na.data.frame(dframe) == TRUE)

dframe$my_letters <- sort(dframe$my_letters)
print(dframe)

mean(dframe[which(is.na(dframe[,1]) == FALSE),1])

```

