---
title: "Homework6"
author: "Sophie Kogut"
date: "2/19/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---
## Generating Simulated Data
```{r, warning= FALSE, message= FALSE}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation


# quick and dirty, a truncated normal distribution to work on the solution set

z <- rnorm(n=3000,mean=0.2) # 3000 samples, mean of 0.2
z <- data.frame(1:3000,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)

```
### Plot histogram of data
```{r, warning= FALSE, message= FALSE}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
# Histogram; y is rescaled from counts to density, area under the curve = 1

# Add empirical density curve

p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
# Curve is fitted to data, does not assume any particular probability dist but smooths the shape

# Get maximum likelihood parameters for normal

normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute, requires ""

# Plot normal probability density

meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]
# maximum likelihood parameters to be passed to stat_function()
xval <- seq(0,max(z$myVar),len=length(z$myVar))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
 p1 + stat
# normal is symmetric, does not fit this data well because data >0

 # Plot exponential probability density
 
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 p1 + stat + stat2
# exponential also does not fit perfectly
 
# Plot uniform probabilty density
 
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 p1 + stat + stat2 + stat3
# uniform does not fit the data
 
# Plot gamma probability density
 
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
# gamma is best fit so far

# Plot beta probabilty density
 
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
# data must be rescaled between 0-1 so they can be compared to beta, pretty close fit
```


## Using real data
```{r, warning= FALSE, message= FALSE}
# Dataset that includes the difference in R^2 between cases and controls obtained from a kinear model comparing 1 ERV and 1 gene transcript.
z <- read.table("/Users/skogut/Desktop/Thesis/deltaRsq.csv",header=TRUE,sep=",", stringsAsFactors=FALSE)
str(z)
summary(z)

```

### Plot histogram of data
```{r, warning= FALSE, message= FALSE}

p1 <- ggplot(data=z, aes(x=Delta.R.2, y=..density..)) +
  geom_histogram(color="goldenrod4",fill= "royalblue3", size=0.2) 
print(p1)
# Histogram; y is rescaled from counts to density, area under the curve = 1

# Add empirical density curve

p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
# Curve is fitted to data, does not assume any particular probability dist but smooths the shape

# Get maximum likelihood parameters for normal

normPars <- fitdistr(z$Delta.R.2,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute, requires ""

# Plot normal probability density

meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]
# maximum likelihood parameters to be passed to stat_function()
xval <- seq(0,max(z$Delta.R.2),len=length(z$Delta.R.2))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$Delta.R.2), args = list(mean = meanML, sd = sdML))
 p1 + stat
# normal is symmetric, does not fit this data well because data >0

 # Plot exponential probability density
 
expoPars <- fitdistr(z$Delta.R.2,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="green", n = nrow(z), args = list(rate=rateML))
 p1 + stat + stat2
# exponential also does not fit perfectly
 
# Plot uniform probabilty density
 
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="yellow3", n = length(z$Delta.R.2), args = list(min=min(z$Delta.R.2), max=max(z$Delta.R.2)))
 p1 + stat + stat2 + stat3
# uniform does not fit the data
 
# Plot gamma probability density
 
gammaPars <- fitdistr(z$Delta.R.2,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="orchid", n = nrow(z), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4


# Plot beta probabilty density
 
pSpecial <- ggplot(data=z, aes(x=Delta.R.2/(max(Delta.R.2 + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$Delta.R.2/max(z$Delta.R.2 + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$Delta.R.2), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
# data must be rescaled between 0-1 so they can be compared to beta, pretty close fit, do not know why orchid line is so big
```
Using the best-fitting distribution, go back to the code and get the maximum likelihood parameters. Use those to simulate a new data set, with the same length as your original vector, and plot that in a histogram and add the probability density curve. Right below that, generate a fresh histogram plot of the original data, and also include the probability density curve.

### Simulating data: Exponential Distribution
```{r, warning= FALSE, message= FALSE}
# the exponential curve fits this data the best
expoPars <- fitdistr(z$Delta.R.2,"exponential")
rateML <- expoPars$estimate["rate"]
print(expoPars)
sim<- rexp(n= nrow(z),rate = rateML )
sim<- as.data.frame(sim)
str(sim)
psim<- ggplot(data=sim, aes(x=sim, y=..density..)) +
  geom_histogram(color="goldenrod4",fill= "steelblue1", size=0.2) 
print(psim)
xval <- seq(0,max(sim),len=nrow(sim))
statE <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="green", n = nrow(sim), args = list(rate=rateML))
psim + statE
p1+ stat2

```
Yes, my model is doing a good job of simulating the data. The curve fits the data pretty well and the the histogram looks pretty similar. 
